---
title: "Unemployment_Rate"
author: "Raquel Baltazar"
date: "`r format(Sys.time(), '%a %d %b %Y (%H:%M:%S)')`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**This dataset explores the evolution of the total unemployment rate from people between 15 and 74 years old in Europe.**

# Loading Libraries

```{r library}

library(eurostat)
library(knitr)
library(kableExtra)
library(ggplot2)
library(magrittr)
library(data.table)
library(here)
library(countrycode)
library(gganimate)
library(gapminder)
library(data.table)
library(dplyr)

```

```{r query}
query <- search_eurostat(pattern = "Unemployment rate - quarterly data, seasonally adjusted", 
                         type = "table", fixed = FALSE)
query[, 1:2]
```

```{r new_variables}

countries <-  c("AT", "BE", "BG", "CH", "CY", "CZ", "DE", "DK", "EE", "EL", "ES", "FI", "FR", "HR","HU", "IE", "IS", "IT", "LI", "LT","LU", "LV", "MT", "NL", "NO", "PL", "PT","RO", "SE", "SI", "SK", "UK")

PTandEU <-  c("PT","EU27_2020")

```

# Defining variables dat and dat1

## dat = Every country in the EU separately
## dat1 = Portugal vs EU

```{r get}

dat <- get_eurostat(id="tipsun30", time_format = "num", 
                    filters = list (geo = countries))
dat [1:2,]

dat1<- get_eurostat(id="tipsun30", time_format = "num", 
                    filters = list (geo = PTandEU))
                    
dat1 [1:2,]
```

```{r label}
dat <- label_eurostat (dat)
dat[1:3,]


dat1 <- label_eurostat (dat1)
dat1[1:3,]
```

# Clean the data  (remove columns)

```{r clean_dat_dat1}

setDT(dat)

setDT(dat1)

dat[, c("s_adj", "sex", "age", "unit"):=NULL]

dat1[, c("s_adj", "sex", "age", "unit"):=NULL]

setDT(dat)

setDT(dat1)
```

# Evolution of unemployment rate in EU, from 1995 to Q2 of 2020

## Line graph

```{r eurostat_and_plots}

ggplot(dat, 
       aes(x = time, y= values, color = geo, label = geo)) + 
  geom_line (alpha = .5) +
  geom_text (data = dat %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  labs(title = "Unemployment Rate in Europe from Q1 of 1995 to Q2 of 2020",
       x= "Year", y= "%")

  labs(title = 'Year: {frame_time}', x = 'Time', y = 'Percentage') +
  transition_time(time) +
  ease_aes('linear')
```

What about just the last 3 years?

```{r var_last3years}
last3 <- dat [time %in% c("2018.00", "2018.25", "2018.50", "2018.75", "2019.00", "2019.25", "2019.50","2019.75", "2020.00", "2020.25", "2020.50", "2020.75")]
```


```{r eurostat_and_plots1}

ggplot(last3, 
       aes(x = time, y= values, color = geo, label = geo)) + 
  geom_line (alpha = .5) +
  geom_text (data = dat %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  labs(title = "Unemployment Rate in Europe for the last 3 years",
       x= "Year", y= "%")

  labs(title = 'Year: {frame_time}', x = 'Time', y = 'Percentage') +
  transition_time(time) +
  ease_aes('linear')
```

**there is no data for the EU**

# Unemployment Rate in Portugal

```{r eurostat_and_plots2}

ggplot(dat1, 
       aes(x = time, y= values, color = geo, label = geo)) + 
  geom_line (alpha = .5) +
  geom_text (data = dat1 %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  labs(title = "Unemployment Rate in Portugal from Q1 of 1995 to Q3 of 2020",
       x= "Year", y= "%")

  labs(title = 'Year: {frame_time}', x = 'Time', y = 'Percentage') +
  transition_time(time) +
  ease_aes('linear')
```

